ASHA 2023 Listening Effort Data Analysis

### 2023-10-30 ASHA 2023 Analysis/data_analysis.Rmd
Completed dilation reanalysis

# Load Needed Packages
library(rio) # install.packages("rio")
library(tidyverse) # install.packages("tidyverse")
# Need remotes package to install gazer and saccades
library(remotes) # install.packages("remotes")
library(gazer) # remotes::install_github("dmirman/gazer")
library(saccades) # remotes::install_github("tmalsburg/saccades/saccades")
library(zoo) # install.packages("zoo")
library(ggpubr) # installl.packages("ggpubr")
library(gt) # install.packages("gt")
library(glmmTMB) # install.packages("glmmTMB")
library(performance) # install.packages("performance")
library(sjPlot) # install.packages("sjPlot")

@@

@@

Both the conditional and unconditional ICC is 0.99.

@@

To control for trial order, trial_c is added as a fixed effect to model 2. A likelihood ratio test was also conducted to test whether model 2 provides a better fit to the data compared to mdoel 1

@@

Based on the model output, adding speaker to the model improved the model fit to the data (chi-sq(1), p <0.001). The fixed effect estimate indicates that, while controlling for trial order, the perceived listening effort rating across listeners and phrases is about 2.35 higher for ALS speaker.

@@

>>>>

# Load Needed Packages
library(rio) # install.packages("rio")
library(tidyverse) # install.packages("tidyverse")
# Need remotes package to install gazer and saccades
library(remotes) # install.packages("remotes")
library(gazer) # remotes::install_github("dmirman/gazer")
library(saccades) # remotes::install_github("tmalsburg/saccades/saccades")
library(zoo) # install.packages("zoo")
library(ggpubr) # installl.packages("ggpubr")
library(gt) # install.packages("gt")
library(glmmTMB) # install.packages("glmmTMB")
library(performance) # install.packages("performance")
library(sjPlot) # install.packages("sjPlot")
library(ggridges) # install.packages("ggridges")
library(lme4) # install.packages("lme4")
library(lmerTest) # install.packages("lmerTest")

@@

peak_pupil_df %>%
  ggplot() +
  aes(y = speaker,
      x = peak_pupil,
      color = speaker,
      fill = speaker) +
  geom_density_ridges(scale = 0.7, alpha = 0.5) +
  theme_bw() +
  scale_fill_manual(values = my_pal) +
  scale_color_manual(values = my_pal)

@@

Both the conditional and unconditional ICC is 0.099.

@@

To control for trial order, trial_c is added as a fixed effect to model 2. A likelihood ratio test was also conducted to test whether model 2 provides a better fit to the data compared to model 1

@@

Based on the model output, adding speaker to the model improved the model fit to the data (chi-sq(1) = 469.74, p <0.001). The fixed effect estimate indicates that, while controlling for trial order, the perceived listening effort rating across listeners and phrases is about 2.35 higher for ALS speaker.

@@

# Alternative Pupil Dilation Analysis

Since the trials for the ALS speaker are much longer than the control speaker's trials, we are going to conduct an exploratory analysis for the pupil dilation measure. For each of the listener's trial, we are going to extract the peak pupil dilation during the analysis region and model it as a function of speaker.

## Peak Pupil Dilation Descriptives

peak_pupil_df <- pupil_roi %>%
  dplyr::group_by(subject, trial_c, speaker, code) %>%
  dplyr::summarize(peak_pupil = max(pupil.binned)) %>%
  ungroup()

peak_pupil_des <- peak_pupil_df %>%
  dplyr::group_by(speaker) %>%
  dplyr::summarize(pupil = mean(peak_pupil), 
                   sd = sd(peak_pupil, na.rm = T),
                   se = sd/sqrt(n()))

peak_pupil_des %>%
  gt::gt() %>%
  gt::fmt_number(columns = pupil:se,
                 decimals = 2) %>%
  gt::cols_label(speaker = "Speaker",
                  pupil = "Peak Pupil Dilation",
                  sd = "SD",
                  se = "SE")

### Peak Pupil Dilation Distribution Plot

peak_pupil_df %>%
  ggplot() +
  aes(y = speaker,
      x = peak_pupil,
      color = speaker,
      fill = speaker) +
  geom_density_ridges(scale = 0.7, alpha = 0.5) +
  labs(x = "Mean Peak Pupil Dilation (arbitrary units)", y = "Speaker") +
  theme_bw() +
  theme(legend.position = "none", aspect.ratio = 1) +
  scale_fill_manual(values = my_pal) +
  scale_color_manual(values = my_pal)

ggsave("Plots/peak_pupil_dis.png", plot = last_plot(), height = 5, width = 5, units = "in")

```
### Peak Pupil Dilation Bar Plot

peak_pupil_des %>%
  ggplot() +
   aes(x = speaker,
      y = pupil,
      group = speaker,
      color = speaker,
      fill = speaker) +
  geom_bar(stat = "identity", alpha = 0.6) +
  geom_errorbar(aes(x = speaker, ymin = pupil - se, ymax = pupil + se), width = 0.4) +
  labs(x = "Speaker", y = "Mean Peak Pupil Dilation (arbitrary units)") +
  theme_bw() +
  theme(legend.position = "none", aspect.ratio = 1) +
  scale_color_manual(values = my_pal) +
  scale_fill_manual(values = my_pal)

ggsave("Plots/peak_pupil_bar.png", plot = last_plot(), height = 5, width = 5, units = "in")

```
## Analysis

m0 <- lmer(peak_pupil ~ 1 + (1|subject) + (1|code), data = peak_pupil_df)

# Ran into convergence issues, so added the nlminb optimizer to the model.
# Model had simgular fit issues, so removed the (trial_c|code) random effect.

library(optimx)

m1 <- lmer(peak_pupil ~ 1 + (trial_c|subject), 
           data = peak_pupil_df, control = lmerControl(
                           optimizer ='optimx', optCtrl=list(method='nlminb')))

m2 <- lmer(peak_pupil ~ trial_c + (trial_c|subject), 
           data = peak_pupil_df, control = lmerControl(
                           optimizer ='optimx', optCtrl=list(method='nlminb')))

m3 <- lmer(peak_pupil ~ trial_c + speaker + (trial_c|subject) , 
              data = peak_pupil_df, control = lmerControl(
                           optimizer ='optimx', optCtrl=list(method='nlminb')))

m4 <- lmer(peak_pupil ~ trial_c + speaker + trial_c*speaker + (trial_c|subject) , 
              data = peak_pupil_df, control = lmerControl(
                           optimizer ='optimx', optCtrl=list(method='nlminb')))

#### Final Model

final_model <- lmer(peak_pupil ~ trial_c + speaker + (trial_c|subject) , 
              data = peak_pupil_df, control = lmerControl(
                           optimizer ='optimx', optCtrl=list(method='nlminb')))

sjPlot::tab_model(final_model, file = "Final Model Outputs/peak_pupil_model.html")
